Lyman-$\alpha$ photons through rotating outflows

### 2018-03-31 paper/paper.tex
edit

Recent advances in instrumentation have revolutionized the kind of
kinematic studies that can be performed on galaxies with small angular
sizes on the sky.
For instances, spatially resolved spectra on compact dwarf galaxies
are now possible, and they confirm signs of gas kinematics showing
pure rotation kinematics \citep{2015A&A...577A..21C,2017A&A...600A.125C}
Recent mapping of high redshift circumgalactic regions are also able
to resolve kinematic evicend of large scale rotation
\citep{2018MNRAS.473.3907A} 
These systems with star formation, some neutral gas and low dust
contents are the kind of systems that are naturally expected to have
\lya emission.
What is the expected imprint of rotation on a resonant line such as
the \lya line?

The computational \lya modeling of rotating systems has only
been studied until recently by \cite{Garavito14}.
In that work the authors studied the influence of pure solid body
rotation on the \lya line's morpholgy.
They found that rotation indeed introduces changes, the most
noticeable change is the dependence of the spectra with the viewing
angle with respect to the rotation axis.

@@

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{./figures/discussion/doppler}
  \end{center}
  \caption{\textbf{Spectra from receeding/approaching sides of a toy
      model LAE.}  These results correspond to the RT simulation with
    $\vout=25$\kms, $\vrot=50$\kms, $\tauh=10^5$. 
    The spectra were computed for a viewing angle of
    $\theta=90^{\circ}$. 
    \label{fig:doppler}}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{./figures/results/muse_vs_simulations.pdf}
  \end{center}
  \caption{\textbf{Comparison between observations and simulations.}  
    \label{fig:muse}}
\end{figure}

@@

We find that the main effects of rotation on the \lya\ line morphology are:

@@

As an application to observational data our results suggest that
rotation should also be considered as an explanation of the
non-vanishing \lya intensity at the line's center.
We also show how the \lya spectra from two different sides of a galaxy
can detect approaching/receeding gas motions We show that the
distances between the peaks of these two spectra is smaller than the
expected value of $2\times\vrot$ due to the differente weights at the
surface of the emitting regions.  
Finally we also use data from MUSE to demonstrate to what extent the
outputs of the RT results presented here can be expected to match
hundreds of observed \lya spectra.

>>>>

Recent advances in instrumentation have revolutionized the kind of
kinematic studies that can be performed on galaxies with small angular
sizes on the sky.
For instance, spatially resolved spectra on compact dwarf galaxies
have measured clear signs of gas showing pure rotation kinematics
\citep{2015A&A...577A..21C,2017A&A...600A.125C}; recent mapping of
high redshift circumgalactic regions have also revealed kinematic
evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
These systems with star formation, some neutral gas and low dust
contents naturally show a \lya emission line \citep{PartridgePeebles},
motivating the observational work to phenomenologically link
kinematics traced by H$\alpha$ to \lya spectra \citep[e.g.][]{Herenz2016}.

Theoretically speaking, what is then the expected imprint of rotation
on a resonant emission line such as the \lya line? To what extent is
possible to constrain rotational kinematics from the \lya emission line? 
Detailed radiative transfer (RT) \lya modeling of rotating systems
started until recently by \cite{Garavito14}.
In that work the authors studied the influence of pure solid body
rotation on the \lya line's morphology.
They found that rotation indeed introduces changes, the most
noticeable being the dependence of the spectra with the viewing
angle with respect to the rotation axis.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{./figures/discussion/doppler}
  \end{center}
  \caption{\textbf{Spectra from receeding/approaching sides of a toy
      model LAE.}  These results correspond to the RT simulation with
    $\vout=25$\kms, $\vrot=50$\kms, $\tauh=10^5$. 
    The spectra were computed for a viewing angle of
    $\theta=90^{\circ}$. This toy model illustrates to what extent spectra from
    opposite sides of a galaxy have an imprint of the rotational
    kinematics. 
    \label{fig:doppler}}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{./figures/results/muse_vs_simulations.pdf}
  \end{center}
  \caption{\textbf{Comparison between simulations and observations.}  
    \label{fig:muse}}
\end{figure}

@@

@@

To address the first questionswe posed in the introduction (What is then
the expected imprint of rotation on a resonant emission line such as
the \lya line?)

We can saythat the main effects of rotation on the \lya\ line morphology are:

@@

Addressing the second question we posed in the introduction (To what
extent is possible to constrain rotational kinematics from the \lya
emission line?)  we can say that as an application to observational data our results suggest that
rotation should also be considered as an explanation of the
non-vanishing \lya intensity at the line's center.

A more straightfoward approach is takes into account how the
\lya spectra from two different sides of a galaxy can detect
approaching/receeding gas motions.
In that case one has to be careful and take into account that the
peaks of these two spectra is smaller than the expected value of
$2\times\vrot$ due to the differente weights at the surface of the
emitting regions.

We also use data from MUSE to demonstrate to what extent the
outputs of the RT results presented here can be expected to match
hundreds of observed \lya spectra, but a full constraint of rotation
properties certainly require a full Markov Chain Monte Carlo
exploration of parameter space.
